---
title: "Classes in `hpiR`"
author: "Andy Krause"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
  knitr::opts_chunk$set(echo = TRUE, collapse=TRUE, messages=FALSE, eval=TRUE)
```

## Introduction

This vignette explains the various classes of objects and data used in the `hpiR` package.  It also (re)creates the example data objects that are attached to the package.  

## Class Structures

The table below shows the basic information on the various S3 classes used in the package.  The naming convention of classes is such that prefixes indicate the domain (or the type) of the information contained with: 

* **"series"** = objects refering to information relevant to the entire hpi series (a set of hpi collections)
* **"hpi"** = objects refering to information relevant to the entire hpi collection
* **"index"** = objects refering to information relevant only to the index itself
* **"rt"** or **"hed"** =  objects with information specific to a particular index method ("rt" for repeat transaction and "hed" for hedonic model)^[Additional methods will be added here in the future.]
    * Suffixes of these classes currently include **"data"** and **"model"**
* **"plot"** = objects containing plot objects.  

The `hpi` collection object is the "core" unit of the `hpiR` package. The object will contain all the information, temporary object and analysis about a single house price index.  

The `hpiseries` object is a "super-unit" that contains multiple house price indexes within it.  Indexes within a series share the same data, modeling approach and parameters; they differ only in the length of the indexes.  Additionally there are two types of analyses that can only be applied to series: forecast accuracy (`calcForecastError`) and revision (`calcRevision`). Also note that in an `hpiseries` object the data objects are moved out of the `hpi` objects and into the `hpiseries` object to avoid duplication of data storage.  

| Type | Class | Inherits | Created by | Description |
| :---- | :----- | :------: | :-----: | :-------------------------------------- | 
| **Super** | | | | |
| | `serieshpi` | *~many* | `createSeries()` | A series of progressively longer indexes.  
||||||
| **Core** | | | | |
| | `hpi` | *~many* | ** *****Index()** | Unified house price index object (S3 list) containing data, model and index as well as additional analytics that can be added to it.  
| | `hpiaccuracy` | `data.frame` | **calcAccuracy()**, and subordinates | Calculate the error |
||||||
| **Data** | | | | |
| | `hpidata` | `data.frame` | **dateToPeriod()** | Transaction data.frame with dates converted to relative periods |
| | `rtdata` | `hpidata`, `data.frame` | **rtCreateTrans()** | Data.frame of repeat transactions with standardized field names |
| | `heddata` | `hpidata`, `data.frame` | **hedCreatTrans()** | Data.frame of transactions with some standardized field names for use in hedonic model approach |
||||||
| **Model** | | | | |
| | `hpimodel` | Multiple (ex. `lm`) | **hpiModel()** | Flexible list containing results of the particular model selected |
| | `rtmodel` | Multiple (ex. `lm`) | **rtModel()** | List containing model results of repeat transaction model
| | `hedmodel` | Multiple (ex. `lm`) | **hedModel()** | List containing model results of hedonic model
||||||
| **Index** | | | | |
||||||
| | `hpiindex` | ~`ts` | **modelToIndex()** | List containing index information.  `index` slot is a `ts` object |
| | `indexsmooth` | `ts` | **smoothIndex()** | smooths an index based on the order argument | 
| | `indexvolatility` | NONE | **calcVolatility()** | contains rolling, mean and median volatility of an index | 
||||||
| **Series** | | | | |
| | `seriesrevision` | NONE | **calcRevision()** | Calculate the period, mean and median revision |
| | `seriesaccuracy` | `hpiaccuracy` | **calcSeriesAccuracy()** | Calculate the period, mean and median revision |
||||||
| **Visualization** | | | | |
| | `plotindex` | `gg`, `ggplot` | **plot.hpiindex()** | Plotting object for an index |
| | `plotseries` | `gg`, `ggplot` | **plot.serieshpi()** | Plotting object for an index series |
| | `plotrevision` | `gg`, `ggplot` | **plot.seriesrevision()** | Plotting object for revision statistics |
| | `plotaccuracy` | `gtable`, `gTree`, `grob`, `gDesc` | **plot.hpiaccuracy()** | Plotting object for a error results|
| | `plotvolatility` | `gg`, `ggplot` | **plot.indexvolatility()** | Plotting object for an index |

## Examples And Explanation

Small examples of any of the classes in the table above are available in the `hpiR` as pre-loaded data.  Use `data(ex_'name of class')` to access. Ex: `data(ex_hpidata)`. 

### Loading

Begin by loading the `hpiR` package. We'll assume that you've followed the instructions in the **"introduction"** vignette in order to obtain the package.  

```{r silent_library_load, echo=FALSE}

  suppressMessages(suppressWarnings(library(hpiR)))
  suppressMessages(suppressWarnings(library(knitr)))

```

```{r hpir_load}
  
  library(hpiR)

```

Next, load the example dataset of Seattle house transactions into your R session.  We'll limit to just the central city (**area** equal to 13, 14, and 15).  This can also be accessed via `data(ex_sales)`.

```{r data_load}

  # Load all sales
  data(seattle_sales)

  # Create a smaller sample of sales
  ex_sales <- seattle_sales %>%
    dplyr::filter(area %in% c(13, 14, 15))

  # Or just load it
  data(ex_sales)
  
```

### Data

#### `hpidata` objects

`hpidata` objects inherit from `data.frame`, but are very flexible and the collection of fields that may be in these objects is solely dependent on the data being used.  The only requirements for creating an `hpidata` object is that there must be a date field that can be turned into a period field.  

The `dateToPeriod()` function will create an `hpidata` object by converting the date into a period.  The only difference between the `hpidata` object and the original data.frame is the **trans_date** and  **trans_period** columns and the **periodicity**, **min_date**, **max_date** and **period_table** attributes. 

```{r create_hpidata}

  ex_hpidata <- dateToPeriod(trans_df = ex_sales,
                             date = 'sale_date',
                             periodicity = 'monthly')

```

```{r class_hpidata}

  class(ex_hpidata)

```

```{r str_hpidata}

  str(ex_hpidata)
  
```

#### Method-based data objects

Each of the methods (currently just **rt** and **hed**) will have their own data object classes that include the necessary fields and structures for these method types.  The name format here is "xxxdata"", where 'xxx' is the method abbreviation.  

The `rtdata` object holds repeat transactions and contains the eight standard fields shown below as well as the **period_table** attribute.  

```{r create_rtdata}

  ex_rtdata <- rtCreateTrans(trans_df = ex_hpidata,
                             prop_id = 'pinx',
                             trans_id = 'sale_id',
                             price = 'sale_price')
```

```{r class_rtdata}

  class(ex_rtdata)

```

```{r str_rtdata}

  str(ex_rtdata)
  
```

The `heddata` object requires no changes from the original `hpidata` object and looks the same, with the exception of field standardization (**pinx** is now **prop_id**, **sale_id** is now **trans_id**, **sale_price** is now **price**, **sale_date** is now **date**. ) 

```{r create_heddata}

  ex_heddata <- hedCreateTrans(trans_df = ex_sales,
                               prop_id = 'pinx',
                               trans_id = 'sale_id',
                               price = 'sale_price',
                               date = 'sale_date',
                               periodicity = 'monthly')

```

```{r class_heddata}

  class(ex_heddata)

```

```{r str_heddata}
  
    str(ex_heddata)

```

### Modeling 

Results from modeling the data prior to creating indexes is stored in `hpimodel` objects.  These object contain eight standard slots, shown below.  The slots are standardized across methods, with the exception of the `model_obj` slot which will contain an object that is specific to the method being used to estimate the index. 

```{r create_timematrix, echo=FALSE, eval=FALSE}

  ex_timematrix <- rtTimeMatrix(ex_rtdata)

```    


```{r create_hpimodel}
  
  ex_hpimodel <- hpiModel(hpi_df = ex_rtdata,
                          estimator = 'base',
                          log_dep = TRUE)

```

```{r class_hpimodel}

   class(ex_hpimodel)

```

```{r str_hpimodel}

  str(ex_hpimodel, max.level=1)

```    



Below we see an example of a `model_obj` from a repeat transactions model. This is very similar to what you'd get from a call to a linear model (`lm(...)`)

```{r create_model_obj}
  ex_rtmodel <- ex_hpimodel$model_obj

```

```{r class_model_obj}
  class(ex_rtmodel)

```

```{r str_model_obj}
  str(ex_rtmodel, max.level=1, give.attr=FALSE)

```

We also show an example of a `model_obj` from a hedonic price method, again similar to an `lm(...)` output. 

```{r create_hmodel_obj}
  ex_hedmodel <- hpiModel(hpi_df = ex_heddata,
                          estimator = 'base',
                          dep_var = 'price',
                          ind_var = c('tot_sf', 'beds', 'baths'),
                          log_dep = TRUE)$model_obj

```

```{r class_hmodel_obj}
  class(ex_hedmodel)

```
  
```{r str_hmodel_obj}
    str(ex_hedmodel, max.level=1, give.attr=FALSE)

```

### Index and Smoothing

The `modelToIndex` function will take an `hpimodel` object and turn it into an `hpiindex` object.  `hpiindex` object are fairly basic, contain five standard slots. 

```{r create_hpiindex}
  ex_hpiindex <- modelToIndex(model_obj = ex_hpimodel)

```

```{r class_hpiindex}
  class(ex_hpiindex)

```

```{r str_hpiindex}  
  str(ex_hpiindex, max.level=1)
  
```

If we want to add a smoothed index to this object, we use the `smoothIndex` function.  In its default state, this function just return a simple `ts` (time-series) object of the smoothed index. 

```{r create_smoothindex}
  ex_smoothindex <- smoothIndex(ex_hpiindex)

```

```{r class_smoothindex}
  class(ex_smoothindex)

```

```{r str_smoothindex}
  str(ex_smoothindex)

```

To add it to the `hpiindex` object, we add the **"in_place = TRUE"** argument which will add it to the **smooth** slot in the `hpiindex` object. Note that the class of **temp_index** stays the same, only teh **smooth** slot is added with an object of class `smoothindex` (inheriting from `ts`).

```{r smooth_inplace}
  temp_index <- smoothIndex(ex_hpiindex, in_place = TRUE)

```

```{r class_smooth_inplace}
  class(temp_index)
  
```

```{r class_smooth_inplaces}
    class(temp_index$smooth)
    
```

```{r str_smooth_inplace}
    str(temp_index)

```

If you want to store plots of `hpiindex` objects, the are returned as `plotindex` objects, inheriting from `gg` and `ggplot` (from the `ggplot2` package)

```{r create_plotindex}
  ex_plotindex <- plot(ex_hpiindex)

```

```{r class_plotindex}
  class(ex_plotindex)

```

### Wrapper

The `rtIndex` (and associated `hedIndex`) wrapper combine the above steps -- data prep, modeling and index creation -- into a single function, the output of which is an hpi collection, or simply, an `hpi` object.  This objects is combination of a data object (`hpidata`), a model object (`hpimodel`) and an index object (`hpiindex`).  

```{r create_hpi}
  ex_hpi <- rtIndex(trans_df = ex_rtdata,
                    estimator = 'robust',
                    log_dep = TRUE,
                    trim_model = TRUE,
                    smooth = TRUE)

```

```{r class_hpi}
  class(ex_hpi)

```

```{r str_hpi}
  str(ex_hpi, max.level=1)
 
```

### Analyzing Indexes

#### Volatility

Index volatility can be calculated by providing an `hpiindex` object to the `calcVolatility()` function. The result is an object of class `indexvolatility`, containing three vector objects as well as two attributes indicating the original index and the window used. 

```{r create_indexvolatility}
  ex_indexvolatility <- calcVolatility(index = ex_hpiindex,
                                       window = 5)

```

```{r class_indexvolatility}
  class(ex_indexvolatility)

```

```{r str_indexvolatility}
  str(ex_indexvolatility)
  
```

The `indexvolatility` object can be nested within an `hpiindex` object by setting **in_place = TRUE**. 

```{r vol_inplace}
  temp_index <- calcVolatility(index = ex_hpiindex,
                               window = 3, 
                               in_place = TRUE)

```

```{r class_volinplace1}
  class(temp_index)

```

```{r class_volinplace2}
  class(temp_index$volatility)

```

```{r str_volinplace}
  str(temp_index, max.level=1, give.attr=FALSE)

```

Volatility plots can be saved in objects of class `plotvolatility`, inheriting from `gg` and `ggplot`

```{r create_volplot}
  ex_plotvolatility <- plot(ex_indexvolatility)

```

```{r class_volplot}
  class(ex_plotvolatility)
 
```

#### Accuracy (Fitting)

Calculating an indexes (fitting) accuracy through either **insample** or **kfold** (out of sample) methods via calcAccuracy returns an object of class `hpiaccuracy`, inheriting from `data.frame`.  These object contain four fields, shown below.  In `hpiaccuracy` objects, each observation in that is being tested is only found once.  This differs from the `seriesaccuracy` objects which are discussed later in which an observation may be tested against an index for each index in the series.  

```{r create_hpiaccr}
  ex_hpiaccuracy <- calcAccuracy(hpi_obj = ex_hpi,
                                 test_type = 'rt',
                                 test_method = 'insample')
```

```{r class_hpiaccr}
  class(ex_hpiaccuracy)

```

```{r str_hpiaccr}
  str(ex_hpiaccuracy)
  
```

`hpiaccuracy` object can be added directly to an `hpi` object with the **in_place = TRUE** argument.^[Accuracy results are added to `hpi` objects and not `hpindex` object because they require data that is outside of the `hpiindex` object and therefore relate to the hpi collection more than just the index.] 

```{r hpia_inplace}
  temp_hpi <- calcAccuracy(hpi_obj = ex_hpi,
                           test_type = 'rt',
                           test_method = 'insample',
                           in_place = TRUE)
```

```{r class_hpia_inplace1}
  class(temp_hpi)

```

```{r class_hpia_inplace2}
  class(temp_hpi$accuracy)

```

```{r str_hpia_inplace}
  str(temp_hpi, max.level=1, give.attr=FALSE)

```

An `hpiaccuracy` object can be plotted and save to a `plotaccuracy` object, but need the `return_plot=TRUE` argument to return to an object. 

```{r create_plotaccr}
  ex_plotaccuracy <- plot(ex_hpiaccuracy, return_plot = TRUE)

```

```{r class_plotaccr}
  class(ex_plotaccuracy)
 
```

#### Series

A series of hpis are created with the `createSeries` function.  This results in a `serieshpi` object, with two initial slots: **data** (`hpidata` object) and **hpis** a list of `hpi` objects (minus the data component which has been removed to avoid duplication.)  

```{r create_serieshpi}

  ex_serieshpi <- createSeries(hpi_obj = ex_hpi,
                               train_period = 48,
                               max_period = 60)
```

```{r class_serieshpi}
  class(ex_serieshpi)

```

```{r str_serieshpi}
  str(ex_serieshpi, max.level=1)
 
```

Series can be plotted and saved into a `plotseries` object. 

```{r create_plotserieshpi}
  ex_plotseries <- plot(ex_serieshpi)

```

```{r class_plotserieshpi}
  class(ex_plotseries)
 
```

#### Revision

The revision of a series is calculated with the `calcRevision()` function and return an object of class `seriesrevision`.  This object contains a data.frame of revision by period and median and mean summaries of overall revision levels. 

```{r create_seriesrevision}
  ex_seriesrevision <- calcRevision(series_obj = ex_serieshpi)

```

```{r class_seriesrevision}
  class(ex_seriesrevision)
  
```

```{r str_seriesrevision}  
  str(ex_seriesrevision)
 
```

To add the revision to the `serieshpi` object, use **in_place = TRUE**.  

```{r srev_inplace}
  temp_series <- calcRevision(series_obj = ex_serieshpi,
                              in_place = TRUE)

```

```{r class_srev_inplace1}
  class(temp_series)

```

```{r class_srev_inplace2}
  class(temp_series$revision)

```

```{r str_srev_inplace}
  str(temp_series, max.level=1)

```

Plotting revision can be saved to a `plotrevision` object.

```{r create_plotrevision}
  ex_plotrevision <- plot(ex_seriesrevision)

```

```{r class_plotrevision}
  class(ex_plotrevision)
 
```

### Series Accuracy

Finally, accuracy can be calculated the series level by providing a `serieshpi` object to the `calcSeriesAccuracy()`, resulting in a `seriesaccuracy` object.  `seriesaccuracy` objects look nearly identical to `hpiaccuracy` objects, however they differ in two respects: 1) they may include estimate of accuracy for the same observation multiple times ^[This will not happen only when using a test_method of type "forecast" and a "forecast_length" of 1.]; 2) they may have a **series** field, indicating from which index in the series the accuracy estimate is from. 

Using the **test_method = 'forecast'** will create a true, prediction-type test of accuracy (out of sample and into the future), where **kfold** is out of sample (but with knowledge of the future) and **insample** is just a simple fitting measure. 

```{r create_seriesaccr}
  ex_seriesaccuracy <- calcSeriesAccuracy(series_obj = ex_serieshpi,
                                          test_method = 'insample',
                                          test_type = 'rt')

```

```{r class_seriesaccr}
  class(ex_seriesaccuracy)

```

```{r str_seriesaccr}
  str(ex_seriesaccuracy)

```  
  
This object can be added directly to the `serieshpi` by using the **in_place = TRUE** argument.

```{r create_sainplace}
  temp_series <- calcSeriesAccuracy(series_obj = ex_serieshpi,
                                    test_method = 'insample',
                                    test_type = 'rt',
                                    in_place=TRUE)

```

```{r class_sainplace}
  class(temp_series)

```

```{r class_sainplace2}
  class(temp_series$accuracy)

```

```{r str_sainplace}
  str(temp_series, max.level=1)
  
``` 

Finally, `seriesaccuracy` objects can be saved to a plot of class `plotaccuracy`

```{r create_plotsaccuracy}
  ex_plotaccuracy <- plot(ex_seriesaccuracy, return_plot = TRUE)

```

```{r class_plotsaccuracy}
  class(ex_plotaccuracy)
 
```

```{r create_package_data_objects, echo=FALSE, eval=FALSE}
  
  # Code for saving objects to the package /data.  Not executed or shown. 

  devtools::use_data(ex_sales, overwrite=TRUE)

  devtools::use_data(ex_hpidata, overwrite=TRUE)
  devtools::use_data(ex_rtdata, overwrite=TRUE)
  devtools::use_data(ex_heddata, overwrite=TRUE)
  
  devtools::use_data(ex_hpimodel, overwrite=TRUE)
  devtools::use_data(ex_timematrix, overwrite=TRUE)
  devtools::use_data(ex_rtmodel, overwrite=TRUE)
  devtools::use_data(ex_hedmodel, overwrite=TRUE)
  
  devtools::use_data(ex_hpiindex, overwrite=TRUE)
  devtools::use_data(ex_smoothindex, overwrite=TRUE)
  devtools::use_data(ex_plotindex, overwrite=TRUE)
  
  devtools::use_data(ex_hpi, overwrite=TRUE)
  
  devtools::use_data(ex_indexvolatility, overwrite=TRUE)
  devtools::use_data(ex_plotvolatility, overwrite=TRUE)
  
  devtools::use_data(ex_hpiaccuracy, overwrite=TRUE)  
  devtools::use_data(ex_plotaccuracy, overwrite=TRUE)  
  
  devtools::use_data(ex_serieshpi, overwrite=TRUE)
  devtools::use_data(ex_plotseries, overwrite=TRUE)
  
  devtools::use_data(ex_seriesaccuracy, overwrite=TRUE)
  
  devtools::use_data(ex_seriesrevision, overwrite=TRUE)
  devtools::use_data(ex_plotrevision, overwrite=TRUE)
  
```
