---
title: "Classes in `hpiR`"
author: "Andy Krause"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

## Introduction

This vignette explains the various classes of objects and data used in the `hpiR` package.  It also (re)creates the example data objects that are attached to the package.  

## Class Structures

The table below shows the basic information on the various S3 classes used in the package.  The naming convention of classes is such that prefixes indicate the domain of the information contained with: 

* **"super"** = objects refering to information relevant to the entire hpi series
* **"hpi"** = objects refering to information relevant to the entire hpi collection
* **"index"** = objects refering to information relevant only to the index itself
* **"rt"** or **"hed"** =  objects with information specific to a particular index method ("rt" for repeat transaction and "hed" for hedonic model)^[Additional methods will be added here in the future.]
    * Suffixes of these classes include **"data"** and **"model"**
* **"plot"** = objects containing plot objects.  

The `hpi` collection object is the "core" unit of the `hpiR` package. The object will contain all the information, temporary object and analysis about a single house price index.  

The `hpiseries` object is a "super-unit" that contains multiple house price indexes within it.  Indexes within a series share the same data, modeling approach and parameters; they differ only in the length of the indexes.  Additionally there are two types of analyses that can only be applied to series: forecast accuracy (`calcForecastError`) and revision (`calcRevision`). Also note that in an `hpiseries` object the data objects are moved out of the `hpi` objects and into the `hpiseries` object to avoid duplication of data storage.  

| Type | Class | Inherits | Created by | Description |
| :---- | :----- | :------: | :-----: | :-------------------------------------- | 
| **Super** | | | | |
| | `serieshpi` | *~many* | `createSeries()` | A series of progressively longer indexes.  
||||||
| **Core** | | | | |
| | `hpi` | *~many* | ** *****Index()** | Unified house price index object (S3 list) containing data, model and index as well as additional analytics that can be added to it.  
||||||
| **Data** | | | | |
| | `hpidata` | `data.frame` | **dateToPeriod()** | Transaction data.frame with dates converted to relative periods |
| | `rtdata` | `hpidata`, `data.frame` | **rtCreateTrans()** | Data.frame of repeat transactions with standardized field names |
| | `heddata` | `hpidata`, `data.frame` | **hedCreatTrans()** | Data.frame of transactions with some standardized field names for use in hedonic model approach |
||||||
| **Model** | | | | |
| | `hpimodel` | Multiple (ex. `lm`) | **hpiModel()** | Flexible list containing results of the particular model selected |
| | `rtmodel` | Multiple (ex. `lm`) | **rtModel()** | List containing model results of repeat transaction model
| | `hedmodel` | Multiple (ex. `lm`) | **hedModel()** | List containing model results of hedonic model
||||||
| **Index** | | | | |
| | `hpiindex` | ~`ts` | **modelToIndex()** | List containing index information.  `index` slot is a `ts` object |
||||||
| | `indexsmooth` | `ts` | **smoothIndex()** | smooths an index based on the order argument | 
| | `indexvolatility` | NONE | **calcVolatility()** | contains rolling, mean and median volatility of an index | 
| | `indexaccuracy` | `data.frame` | **calcAccuracy()**, and subordinates | Calculate the error |
||||||
| **Series** | | | | |
| | `seriesrevision` | NONE | **calcRevision()** | Calculate the period, mean and median revision |

||||||
| **Visualization** | | | | |
| | `plotindex` | `gg`, `ggplot` | **plot.hpiindex()** | Plotting object for an index |
| | `plotseries` | `gg`, `ggplot` | **plot.serieshpi()** | Plotting object for an index series |
| | `plotrevision` | `gg`, `ggplot` | **plot.seriesrevision()** | Plotting object for revision statistics |
| | `plotaccuracy` | `gtable`, `gTree`, `grob`, `gDesc` | **plot.indexerrors()** | Plotting object for a error results|



| | `volatiltyplot` | `gg`, `ggplot` | **plot.hpiindex()** | Plotting object for an index |

## Examples

Begin by loading the `hpiR` package. We'll assume that you've followed the instructions in the **"introduction"** vignette in order to obtain the package.  

```{r eval=TRUE, messages=FALSE, echo=FALSE}

  suppressMessages(suppressWarnings(library(hpiR)))
  suppressMessages(suppressWarnings(library(knitr)))

```

```{r eval=TRUE, messages=FALSE, echo=TRUE}
  
  library(hpiR)

```

Next, load the example dataset of Seattle house transactions into your R session.  We'll limit to just the central city (**area** equal to 13, 14, 15 and 79)

```{r eval=TRUE, messages=FALSE, echo=TRUE}

  data(seattle_sales)

  central_sales <- seattle_sales %>%
    dplyr::filter(area %in% c(13, 14, 15, 79))

```

#### Data

An example `hpidata` object

```{r eval=TRUE, messages=FALSE, echo=TRUE}

  ex_hpidata <- dateToPeriod(trans_df = central_sales,
                             date = 'sale_date',
                             periodicity = 'monthly')
  class(ex_hpidata)

```

An example `rtdata` object

```{r eval=TRUE, messages=FALSE, echo=TRUE}

  ex_rtdata <- rtCreateTrans(trans_df = ex_hpidata,
                             prop_id = 'pinx',
                             trans_id = 'sale_id',
                             price = 'sale_price')
  class(ex_rtdata)

```

```{r eval=TRUE, messages=FALSE, echo=TRUE}

  ex_heddata <- hedCreateTrans(trans_df = central_sales,
                               prop_id = 'pinx',
                               trans_id = 'sale_id',
                               price = 'sale_price',
                               date= 'sale_date',
                               periodicity = 'monthly')
  class(ex_heddata)

```

### Modeling 

```{r eval=TRUE, messages=FALSE, echo=TRUE}

  ex_hpimodel <- hpiModel(hpi_df = ex_rtdata,
                          estimator = 'base',
                          log_dep = TRUE)

  ex_rtmodel <- ex_hpimodel$model_obj
  
  class(ex_hpimodel)
  class(ex_rtmodel)

```

```{r eval=TRUE, messages=FALSE, echo=FALSE}

  ex_hpimodel_h <- hpiModel(hpi_df = ex_heddata,
                            estimator = 'base',
                            dep_var = 'price',
                            ind_var = c('tot_sf', 'beds', 'baths'),
                            log_dep = TRUE)
  ex_hedmodel <- ex_hpimodel_h$model_obj

  class(ex_hpimodel_h)
  class(ex_hedmodel)

```

### Index and Smoothing

```{r eval=TRUE, messages=FALSE, echo=FALSE}

  ex_hpiindex <- modelToIndex(model_obj = ex_hpimodel)
  class(ex_hpiindex)
  
```

```{r eval=TRUE, messages=FALSE, echo=FALSE}

  ex_plotindex <- plot(ex_hpiindex)
  class(ex_plotindex)

```

```{r eval=TRUE, messages=FALSE, echo=FALSE}

  ex_smoothindex <- smoothIndex(ex_hpiindex)
  class(ex_smoothindex)

```

### Wrapper

```{r eval=TRUE, messages=FALSE, echo=FALSE}

  ex_hpi <- rtIndex(trans_df = ex_rtdata,
                    estimator = 'robust',
                    log_dep = TRUE,
                    trim_model = TRUE,
                    smooth = TRUE)
  
  class(ex_hpi)
 
```

### Analysis

```{r eval=TRUE, messages=FALSE, echo=FALSE}

  ex_indexvolatility <- calcVolatility(index = ex_hpiindex,
                                       window = 5)
  
  class(ex_indexvolatility)
 
```

```{r eval=TRUE, messages=FALSE, echo=FALSE}

  ex_plotvolatility <- plot(ex_indexvolatility)
  class(ex_plotvolatility)
 
```

```{r eval=TRUE, messages=FALSE, echo=FALSE}

  ex_indexaccuracy <- calcAccuracy(hpi_obj = ex_hpi,
                                   test_type = 'rt',
                                   test_method = 'insample')
  
  class(ex_indexaccuracy)
 
```

```{r eval=TRUE, messages=FALSE, echo=FALSE}

  ex_plotaccuracy <- plot(ex_indexaccuracy, return_plot = TRUE)
  class(ex_plotaccuracy)
 
```

```{r eval=TRUE, messages=FALSE, echo=FALSE}

  ex_serieshpi <- createSeries(hpi_obj = ex_hpi,
                               train_period = 48,
                               max_period = 60)
  class(ex_serieshpi)
 
```

```{r eval=TRUE, messages=FALSE, echo=FALSE}

  ex_plotseries <- plot(ex_serieshpi)
  class(ex_plotseries)
 
```


```{r eval=TRUE, messages=FALSE, echo=FALSE}

  ex_seriesrevision <- calcRevision(series_obj = ex_serieshpi)
  class(ex_seriesrevision)
 
```

```{r eval=TRUE, messages=FALSE, echo=FALSE}

  ex_plotrevision <- plot(ex_seriesrevision)
  class(ex_plotrevision)
 
```

#### Saving objects to the package

```{r eval=TRUE, messages=FALSE, echo=FALSE}
  
  devtools::use_data(ex_hpidata, overwrite=TRUE)
  devtools::use_data(ex_rtdata, overwrite=TRUE)
  devtools::use_data(ex_heddata, overwrite=TRUE)
  devtools::use_data(ex_hpimodel, overwrite=TRUE)
  devtools::use_data(ex_rtmodel, overwrite=TRUE)
  devtools::use_data(ex_hedmodel, overwrite=TRUE)
  devtools::use_data(ex_hpiindex, overwrite=TRUE)
  devtools::use_data(ex_smoothindex, overwrite=TRUE)
  devtools::use_data(ex_plotindex, overwrite=TRUE)
  
  devtools::use_data(ex_hpi, overwrite=TRUE)
  devtools::use_data(ex_indexvolatility, overwrite=TRUE)
  devtools::use_data(ex_plotvolatility, overwrite=TRUE)
  
  devtools::use_data(ex_indexaccuracy, overwrite=TRUE)  
  devtools::use_data(ex_plotaccuracy, overwrite=TRUE)  
  
  devtools::use_data(ex_hpiseries, overwrite=TRUE)
  devtools::use_data(ex_plotseries, overwrite=TRUE)
  
  devtools::use_data(ex_seriesrevision, overwrite=TRUE)
  devtools::use_data(ex_plotrevision, overwrite=TRUE)
  
  

```
